Radio Emission from Red-Giant Hot Jupiters

### 2015-11-23 ms.tex
math corrected, minor edits

\documentclass[iop,numberedappendix,apj,]{emulateapj}

@@

The radius of the magnetic stand-off point, $r_{\rm mag}$, is obtained based on the balance between the stellar wind pressure and the planetary magnetic pressure: 
%%%
\begin{equation}
m_p n v ^2 \sim \frac{B^2}{8\pi}\left( \frac{r_{\rm mag}}{R_p} \right)^{-6}  \label{eq:stand-off} ,
\end{equation}
%%%
where $m_p$ is the proton mass. 
\revise{Thus, }{In equation (\ref{eq:stand-off}), we ignore the ram pressure of the particles from the planets. %
}
%%%
\begin{equation}
r_{\rm mag} \sim R_p \left( 8\pi m_p n \right) ^{-1/6} v^{-1/3} B^{1/3} \label{eq:stand-off-radius} 
\end{equation}
%%%
The radius obtained for Jupiter from this equality is about half of the actual magnetospheric radius \citep[][]{griesmeier2005}. 
To estimate $r_{\rm mag}$ for RGHJs, we scale this radius according to the parameter dependence of equation~(\ref{eq:stand-off-radius}).

@@

As discussed in \citet{ignace2010}, as the stars evolve, the ionization fraction of the stellar wind is lowered down to order of $\sim 10^{-3}$ \citep{drake1987}. Since only charged particles interact with planetary magnetic field, this suggests inefficient interaction with planetary magnetosphere and hence low input energy for radio emission. 
However, at this stage, the velocity of the stellar wind becomes slower than the escape velocity of the planetary companion and hence the stellar wind particles are expected to accrete onto the planets. 
\citet{spiegel+madhusudhan2012} \revise{}{considered Bondi–Hoyle accretion and} computed the accretion luminosity $L_{\rm acc}$ and the temperature $T_{\rm acc}$ as follows: 
%%%%%%%%%
\begin{eqnarray}
L_{\rm acc} \sim &&  10^{25} \; {\rm erg\;s}^{-1} \left( \frac{\dot M_*}{10^{-8} M_{\odot }/{\rm yr}} \right)  \notag \\
&& \times \left( \frac{M_p}{M_{\rm J}} \right)^3 \left( \frac{M_*}{M_{\odot }} \right). \\
%
T_{\rm acc} \sim && 2 \times 10^5 ~\mbox{K} \left( \frac{M_p}{M_{\rm J}} \right). \label{eq:Tacc}
\end{eqnarray}
%%%%%%%%%
The accretion onto planets, therefore, leads to emission of UV/X-ray radiation 
whose characteristic energy $\sim k_B T_{\rm acc} $ exceeds the ionization energy of hydrogen, $E_{\rm Rydberg} = 13.6$~eV, or $\lambda = 91.2\rm~nm$. 
Therefore, UV/X-ray radiation from accretion will create local ionized region around the planet.

\revise{}{Note that in this case the accretion radius is}
%%%%%%%%%
\begin{eqnarray}
R_{\rm acc} = \frac{2GM_p}{v_{\rm rel}^2} \sim 
\end{eqnarray}
%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Effect of ram pressure due to planetary outflow}

At the red giant stage, the atmospheric temperature of planets increases to $> 1000 {\rm K}$. At this point, one might anticipate that planetary atmosphere starts to escape. 
Also, the UV/X-ray produced from accretion may input energy for atmospheric molecules to escape. In this section, we briefly estimate these effects on the circumplanetary configuration.

The escape parameter of RGHJs is:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\lambda _c = \frac{G M_p m_{\rm p}}{k T_c r_c} = 214 \left( \frac{M_p}{M_{\rm J}} \right) \left( \frac{T_c}{1000 {\rm K}} \right)^{-1} \left( \frac{r_c}{R_{\rm J}} \right)^{-1}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $m_{\rm p}$ is mass of a hydrogen atom, $r_c$ is the radius of exobase. Therefore in most of the case the escape is safely assumed to be Jeans escape. 
Escape rate of Jeans escape is: 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{equation}
%\dot M = 4 \pi \rho_0 a r^2 \left( \frac{v_{\rm esc}(r_0)}{2a} \right)^2 \exp \left( - \frac{v_{\rm esc}^2 (r_0) }{2 a^2} + \frac{3}{2} \right)
%\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
%\Phi _{\rm Jeans} &=& 4 \pi r_c^2 \cdot \frac{n_c}{2 \sqrt{\pi}} B \sqrt{\frac{2 k_B T_c }{m}} ( 1 + \lambda _c ) e^{-\lambda _c} \\
\Phi _{\rm Jeans} (r_c) &=& \frac{1}{\sigma } \sqrt{ \frac{2 G \bar \rho }{3} } \left( \lambda _c + 1 \right) \sqrt{ \lambda _c } e^{-\lambda _c} \\
%&=& 1/(pi*(0.529e-10)**2)*sqrt(2*6.674e-11*1330./3.0) * (20+1.0)*sqrt(20.)*exp(-20.0) 
%&=& 1/(pi*(0.529)**2)*sqrt(2*66.74*1330./3.0) * (20+1.0)*sqrt(20.)*exp(-20.0) * 1e14
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
and the ram pressure at distance $r$ due to such an outflow is
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
P_{\rm ram} (r) &=& m_{\rm p} \Phi _{\rm Jeans} (r_c) \left( \frac{r}{R_p} \right)^{-2} v_{\rm esc} \\
% 1.674e-27 * 5.3e18 / 100.0 * 59.5e3 
%&=& 5.27 \times 10^{-6} ~{\rm Pa} \; \left( \frac{r}{10~R_p} \right)^{-2} \left( \frac{M_p}{M_{\rm J}} \right)^{1/2}
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The ram pressure of the stellar wind or planetary magnetic pressure (either hand of equation (\ref{eq:stand-off})) is $\sim 2.7 \times 10^{-12} ~{\rm Pa} $. 
This implies that the ram pressure of the planetary outflow would be negligible when $\lambda _c > 15$. [NEED TO CHECK]

>>>>

\documentclass[iop,numberedappendix,apj,twocolappendix,]{emulateapj}

@@

The radius of the magnetic stand-off point, $r_{\rm mag}$, is obtained based on the balance between the stellar wind pressure and the planetary magnetic pressure: 
%%%
\begin{equation}
m_p n v ^2 \sim \frac{B^2}{8\pi}\left( \frac{r_{\rm mag}}{R_p} \right)^{-6}  \label{eq:stand-off} ,
\end{equation}
%%%
where $m_p$ is the proton mass. 
\revise{}{We note that the outflow from heated planets are negligible compared to these pressures, as discussed in Appendix \ref{ap:outflow}. %
}
Thus, 
%%%
\begin{equation}
r_{\rm mag} \sim R_p \left( 8\pi m_p n \right) ^{-1/6} v^{-1/3} B^{1/3} \label{eq:stand-off-radius} 
\end{equation}
%%%
The radius obtained for Jupiter from this equality is about half of the actual magnetospheric radius \citep[][]{griesmeier2005}. 
To estimate $r_{\rm mag}$ for RGHJs, we scale this radius according to the parameter dependence of equation~(\ref{eq:stand-off-radius}).

@@

As discussed in \citet{ignace2010}, as the stars evolve, the ionization fraction of the stellar wind is lowered down to order of $\sim 10^{-3}$ \citep{drake1987}. Since only charged particles interact with planetary magnetic field, this suggests inefficient interaction with planetary magnetosphere and hence low input energy for radio emission. 
However, at this stage, the velocity of the stellar wind becomes slower than the escape velocity of the planetary companion and hence the stellar wind particles are expected to accrete onto the planets. 
\citet{spiegel+madhusudhan2012} \revise{}{considered Bondi--Hoyle accretion where the accretion disk is}
%%%%%%%%%
\begin{eqnarray}
R_{\rm acc} &=& \frac{2GM_p}{v_{\rm rel}^2} \\
&\sim & 4.0~R_{\rm J} \left( \frac{M_p}{M_{\rm J}} \right) \left( \frac{v_{\rm rel}}{30 ~{\rm km~s^{-1}}} \right)^{-2}
\end{eqnarray}
%%%%%%%%%
\revise{}{The accretion luminosity $L_{\rm acc}$ and the temperature $T_{\rm acc}$ are}
%%%%%%%%%
\begin{eqnarray}
L_{\rm acc} \sim &&  10^{25} \; {\rm erg\;s}^{-1} \left( \frac{\dot M_*}{10^{-8} M_{\odot }/{\rm yr}} \right)  \notag \\
&& \times \left( \frac{M_p}{M_{\rm J}} \right)^3 \left( \frac{M_*}{M_{\odot }} \right). \\
%
T_{\rm acc} \sim && 2 \times 10^5 ~\mbox{K} \left( \frac{M_p}{M_{\rm J}} \right). \label{eq:Tacc}
\end{eqnarray}
%%%%%%%%%
The accretion onto planets, therefore, leads to emission of UV/X-ray radiation 
whose characteristic energy $\sim k_B T_{\rm acc} $ exceeds the ionization energy of hydrogen, $E_{\rm Rydberg} = 13.6$~eV, or $\lambda = 91.2\rm~nm$. 
Therefore, UV/X-ray radiation from accretion will create local ionized region around the planet.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Planetary outflow}
\label{ap:outflow}

One might anticipate that, at the red giant stage, planetary atmosphere start to escape due to both the increased temperature due to stellar irradiation (typically $> 1000 {\rm K}$) and the UV/X-ray radiation produced by accretion from stellar wind. 
In this section, we briefly estimate these effects on the circumplanetary configuration.

%One might anticipate that these factors increase the escape of planetary atmospheres, which could affect the pressure balance represented by equation (\ref{eq:stand-off}).

The conventional escape parameter is computed for RGHJs as follows
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\lambda _c = \frac{G M_p m_{\rm p}}{k T_c r_c} = 218 \left( \frac{M_p}{M_{\rm J}} \right) \left( \frac{T_c}{1000 {\rm K}} \right)^{-1} \left( \frac{r_c}{R_{\rm J}} \right)^{-1}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $T_c$ and $r_c$ are the radius and temperature at exobase. In most of the cases we consider, therefore, the escape is safely assumed to be Jeans escape. 
The rate of Jeans escape is 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{equation}
%\dot M = 4 \pi \rho_0 a r^2 \left( \frac{v_{\rm esc}(r_0)}{2a} \right)^2 \exp \left( - \frac{v_{\rm esc}^2 (r_0) }{2 a^2} + \frac{3}{2} \right)
%\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
%\Phi _{\rm Jeans} &=& 4 \pi r_c^2 \cdot \frac{n_c}{2 \sqrt{\pi}} B \sqrt{\frac{2 k_B T_c }{m}} ( 1 + \lambda _c ) e^{-\lambda _c} \\
\Phi _{\rm Jeans} (r_c) &=& \frac{1}{\sigma } \sqrt{ \frac{G M_p}{2 R_p^3} } \left( \lambda _c + 1 \right) \sqrt{ \lambda _c } e^{-\lambda _c} \\
%&=& 1/(pi*(0.529e-10)**2)*sqrt(2*6.674e-11*1330./3.0) * (20+1.0)*sqrt(20.)*exp(-20.0) 
%&=& 1/(pi*(0.529)**2)*sqrt(2*66.74*1330./3.0) * (20+1.0)*sqrt(20.)*exp(-20.0) * 1e14
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $\sigma $ is the cross section of hydrogen atoms, assumed to be equal to Bohr radius here [[OK??]]. 
The ram pressure at distance $r$ due to such an outflow is
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
P_{\rm ram} (r) &=& m_{\rm p} \Phi _{\rm Jeans} (r_c) \left( \frac{r}{R_p} \right)^{-2} v_{\rm esc} 
% 1.674e-27 * 5.3e18 / 100.0 * 59.5e3 
%&=& 5.27 \times 10^{-6} ~{\rm Pa} \; \left( \frac{r}{10~R_p} \right)^{-2} \left( \frac{M_p}{M_{\rm J}} \right)^{1/2}
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
which indicates $3.6\times 10^{-7}~{\rm barye}$ even when $\lambda _c = 1$. 
On the other hand, the ram pressure of the stellar wind or planetary magnetic pressure (either hand of equation (\ref{eq:stand-off})) is $\sim 2.7 \times 10^{-5} ~{\rm barye}$. 
This implies that the ram pressure of the planetary outflow would be negligible.
